The formal verification of the ctm approach to forcing \thanks{Supported by Secyt-UNC project 33620180100465CB and Conicet.}

### 2022-08-26 Independence_CH/main_results.tex
Attempt to clarify the use of axiom schemes in Forcing Theorems

It is to be noted that the \emph{application} of the Forcing Theorems \emph{does not
require any extra replacement}. That is, no further replacement
instance holding at $M$ is necessary to invoke
Theorem~\ref{th:forcing-thms}, independently of the
formula $\phi$. But this seems  not be the case for
Separation, at least from inspecting our formalization: More instances are needed
as the complexity of $\phi$ grows. One point where this is apparent is
in the proof of Theorem~\ref{th:forcing-thms}(\ref{item:truth-lemma}),
that appears as the \isatt{truth{\uscore}lemma} in our development; it
depends on \isatt{truth{\uscore}lemma'} and
\isatt{truth{\uscore}lemma{\uscore}Neg}, which explicitly invoke
\isatt{separation{\uscore}ax}. In any case, our intended grounds
(v.g., the transitive collapse of countable elementary submodels of a
rank initial segment $V_\alpha$ or an $H(\kappa)$) all satisfy full
Separation.

>>>>

It is to be noted that the proofs of the Forcing Theorems do not
require any extra replacement; actually, they only need the 7
instances appearing in \isatt{instances1{\uscore}fms} and
\isatt{instances{\uscore}ground{\uscore}fms}.  But this seems not be
the case for Separation, at least by inspecting our formalization:
More instances holding in $M$ are needed 
as the complexity of $\phi$ grows. One point where this is apparent is
in the proof of Theorem~\ref{th:forcing-thms}(\ref{item:truth-lemma}),
that appears as the \isatt{truth{\uscore}lemma} in our development; it
depends on \isatt{truth{\uscore}lemma'} and
\isatt{truth{\uscore}lemma{\uscore}Neg}, which explicitly invoke
\isatt{separation{\uscore}ax}. In any case, our intended grounds
(v.g., the transitive collapse of countable elementary submodels of a
rank initial segment $V_\alpha$ or an $H(\kappa)$) all satisfy full
Separation.
